GPU-friendly Stroke Expansion

### 2024-06-15 paper.tex
Update GPU timing plots with new strongly correct measurements

\begin{document}
\title{GPU-friendly Stroke Expansion}
\author{Anonymous Author 1}
\author{Anonymous Author 2}
%\affiliation{
%    \institution{Google}
%    \city{San Francisco}
%    \state{CA}
%    \country{USA}
%}
\begin{CCSXML}
    <ccs2012>
       <concept>
           <concept_id>10010147.10010371.10010372</concept_id>
           <concept_desc>Computing methodologies~Rendering</concept_desc>
           <concept_significance>500</concept_significance>
           </concept>
       <concept>
           <concept_id>10010147.10010371.10010396.10010399</concept_id>
           <concept_desc>Computing methodologies~Parametric curve and surface models</concept_desc>
           <concept_significance>500</concept_significance>
           </concept>
     </ccs2012>
\end{CCSXML}

@@

Stroke rendering is an essential and widely part of the standard vector graphics imaging model. The standard representation of paths is a sequence of cubic Bézier segments. While there are a number of published techniques for GPU rendering of filled paths, stroke rendering is more challenging to parallelize, largely because path segments cannot be processed independently of each other; rendering of the \emph{joins} between path segments depends on both adjoining segments, and the endpoints of paths (when not closed) are rendered with \emph{caps}. Joins and caps can be rendered with a variety of styles within the same document. The style of stroke rendering also optionally includes \emph{dashing}, which applies a pattern of dashes and gaps to the outline of the shape, as illustrated in Figure~\ref{fig:stroke_styles}.

@@

There are a number of strategies for rendering strokes. Among them are \emph{local} techniques that break down the stroke into individual closed primitives, distance field techniques (or similar techniques based on point inclusion). In this paper, we focus on \emph{stroke expansion,} the generation of an outline that, when filled, represents the rendered stroke. Implementing stroke rendering using stroke expansion enables a unified approach to path rendering both filled and stroked paths downstream. Among other benefits, such an approach avoids a large number of draw calls when stroked and filled paths are finely interleaved, as is often the case.

@@

Previous work on stroke rendering does not fully solve the problem of fast and correct stroke rendering. Most existing implementations have correctness problems, in particular failing to draw the evolutes and inner joins as required for strong correctness; \citet{Nehab2020} provides a detailed survey. In addition, most implementations are not suitable for running on GPU, requiring sequential execution to produce watertight outlines. One GPU implementation is polar stroking (\citet{Kilgard2020}), a local technique, but among other limitations, it does not bound the Fréchet distance error (rather, it bounds the angle error).

@@

This paper presents a comprehensive solution to the core problem of stroke expansion, well suited to GPU implementation. It can produce both flattened polylines or an approximation consisting of circular arcs. The best choice depends on the capabilities of the path rendering mechanism following stroke expansion, but as we show, outlines consisting of circular arcs have many fewer segments and are correspondingly faster to produce. The algorithm is based on Euler spiral segments as an intermediate representation, with an iterative algorithm based on a straightforward error metric for conversion from cubic Béziers. We have also devised a compact binary encoding of paths, suitable for fully parallel computation of stroke outlines, while requiring minimal CPU-side processing. Our algorithm has been implemented in GPU compute shaders, integrated in a full rendering engine for vector graphics, and shows a dramatic speedup over CPU methods.

@@

Many techniques for stroke expansion described in the literature apply some lowering of cubic Bézier curves to a simpler curve type that is more tractable for evaluating parallel curve. Computing parallel curves directly on cubic Bézier curve segments is not very tractable. In particular, the widely cited Tiller-Hanson algorithm~[\citenum{Tiller1984}] performs well for quadratic Béziers but significantly worse for cubics.

@@

Each thread outputs a polyline that fits a subset of the rendered subpath's outer contour, represented by the path segment (see Figure \ref{fig:stroke_threads}). For a filled primitive, a thread only outputs the flattened approximation of the path segment. For a stroke, a thread outputs the polylines approximating multiple curves: a) the two parallel curves on both sides of the source segment, offset by the desired stroke \emph{half-width}, and b) the joins that connect the path segment to the next adjacent segment, and c) evolute patches in regions of high curvature. We also store metadata in our input encoding that marks the position of a path segment within the containing subpath. We use this information to determine whether to output an end cap instead of a join. We encode one additional segment designated as the \emph{stroke cap marker} for every subpath. The thread assigned to the marker segment only outputs a start cap, which completes the stroke outline.

@@

This procedure is sufficient to generate weakly correct stroke outlines. If the outline contains a cusp -- which can be determined by evaluating the curvature variation along the Euler spiral segment -- we output an evolute patch (as shown in Figure~\ref{fig:evolutes}) in order to achieve strong correctness. A selection of renderings produced by our algorithm that demonstrate both strong and weak correctness is shown in Figure~\ref{fig:stroke_stills}.

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \citet{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in \textasciitilde3 ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \citet{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

\begin{figure}
    \includegraphics[scale=0.73]{test_scenes_gpu_timings}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the stress test scenes. The timings are shown in milliseconds.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from our test scenes}
    \label{fig:test-scenes-gpu-timings}
\end{figure}

>>>>

\begin{document}
\title{GPU-friendly Stroke Expansion}
\author{Raph Levien}
\author{Arman Uguray}
\affiliation{
    \institution{Google}
    \city{San Francisco}
    \state{CA}
    \country{USA}
}
\begin{CCSXML}
    <ccs2012>
       <concept>
           <concept_id>10010147.10010371.10010372</concept_id>
           <concept_desc>Computing methodologies~Rendering</concept_desc>
           <concept_significance>500</concept_significance>
           </concept>
       <concept>
           <concept_id>10010147.10010371.10010396.10010399</concept_id>
           <concept_desc>Computing methodologies~Parametric curve and surface models</concept_desc>
           <concept_significance>500</concept_significance>
           </concept>
     </ccs2012>
\end{CCSXML}

@@

Stroke rendering is an essential part of the standard vector graphics imaging model. The standard representation of paths is a sequence of cubic Bézier segments. While there are a number of published techniques for GPU rendering of filled paths, stroke rendering is more challenging to parallelize, largely because path segments cannot be processed independently of each other; rendering of the \emph{joins} between path segments depends on both adjoining segments, and the endpoints of paths (when not closed) are rendered with \emph{caps}. Joins and caps can be rendered with a variety of styles within the same document. The style of stroke rendering also optionally includes \emph{dashing}, which applies a pattern of dashes and gaps to the outline of the shape, as illustrated in Figure~\ref{fig:stroke_styles}.

@@

There are a number of strategies for rendering strokes. Among them are \emph{local} techniques that break down the stroke into individual closed primitives, distance field techniques (or similar techniques based on point inclusion). In this paper, we focus on \emph{stroke expansion,} the generation of an outline that, when filled, represents the rendered stroke. Implementing stroke rendering using stroke expansion enables a unified approach to rendering both filled and stroked paths downstream. Among other benefits, such an approach avoids a large number of draw calls when stroked and filled paths are finely interleaved, as is often the case.

@@

Previous work on stroke rendering does not fully solve the problem of fast and correct stroking. Most existing implementations have correctness problems, in particular failing to draw the evolutes and inner joins as required for strong correctness; \citet{Nehab2020} provides a detailed survey. In addition, most implementations are not suitable for running on GPU, requiring sequential execution to produce watertight outlines. One GPU implementation is polar stroking (\citet{Kilgard2020}), a local technique, but among other limitations, it does not bound the Fréchet distance error (rather, it bounds the angle error).

@@

This paper presents a solution to the core problem of stroke expansion, well suited to GPU implementation. It can produce both flattened polylines or an approximation consisting of circular arcs. The best choice depends on the capabilities of the path rendering mechanism following stroke expansion, but as we show, outlines consisting of circular arcs have many fewer segments and are correspondingly faster to produce. The algorithm is based on Euler spiral segments as an intermediate representation, with an iterative algorithm based on a straightforward error metric for conversion from cubic Béziers. We have also devised a compact binary encoding of paths, suitable for fully parallel computation of stroke outlines, while requiring minimal CPU-side processing. Our algorithm has been implemented in GPU compute shaders, integrated in a full rendering engine for vector graphics, and shows a dramatic speedup over CPU methods.

@@

Many techniques for stroke expansion described in the literature apply some lowering of cubic Bézier curves to a simpler curve type that is more tractable for evaluating the parallel curve. Computing parallel curves directly on cubic Bézier curve segments is not very tractable. In particular, the widely cited Tiller-Hanson algorithm~[\citenum{Tiller1984}] performs well for quadratic Béziers but significantly worse for cubics.

@@

Each thread outputs a polyline that fits a subset of the rendered subpath's outer contour, represented by the path segment (see Figure \ref{fig:stroke_threads}). For a filled primitive, a thread only outputs the flattened approximation of the path segment. For a stroke, a thread outputs the polylines approximating multiple curves: a) the two parallel curves on both sides of the source segment, offset by the desired stroke \emph{half-width}, b) the joins that connect the path segment to the next adjacent segment, and c) evolute patches in regions of high curvature. We also store metadata in our input encoding that marks the position of a path segment within the containing subpath. We use this information to determine whether to output an end cap instead of a join. We encode one additional segment designated as the \emph{stroke cap marker} for every subpath. The thread assigned to the marker segment only outputs a start cap, which completes the stroke outline.

@@

This procedure is sufficient to generate weakly correct stroke outlines. If the outline contains a cusp, we output an evolute patch (as shown in Figure~\ref{fig:evolutes}) in order to achieve strong correctness. A selection of renderings produced by our algorithm that demonstrate both strong and weak correctness is shown in Figure~\ref{fig:stroke_stills}.

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings_strong}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \citet{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in \textasciitilde3.5 ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \citet{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

\begin{figure}
    \includegraphics[scale=0.73]{test_scenes_gpu_timings_strong}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the stress test scenes. The timings are shown in milliseconds.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from our test scenes}
    \label{fig:test-scenes-gpu-timings}
\end{figure}
